Synthesizing Quotient Lenses

### 2018-02-16 papers/QuotientLenses/main.tex
a few intro edits; checkpoint

Oftentimes, different applications store the same data in custom formats. For
example, \bibtex{} and EndNote both store bibliography data; spreadsheet data
can be represented as a CSV and as a TSV; web apis can return data as a JSON or
XML object. Synchronizing these applications amounts to synchronizing the data
formats. \emph{Lenses} comprise such synchronization functions, and the ``lens
laws'' provide restrictions on how these synchronization functions interoperate.
Bidirectional languages like Boomerang~\cite{boomerang} and
BiGul~\cite{DBLP:conf/pepm/KoZH16} use a single expression to encode all of the
lens's synchronization functions at the same time.

However, these lens laws are too restrictive for many important transformations.
In particular, these lens laws do not permit transformations that put the data
in canonical forms. In \bibtex{}, an author can be specified as \cd{lastname,
firstname} and \cd{firstname lastname}. While it is important to allow authors
to be input in both ways, we would like these authors to only be output as
\cd{firstname lastname}, but the lens laws prevent such functions.
{\em Quotient lenses}~\cite{quotientlenses} are a generalization of lenses where
each of the data formats is quotiented by some equivalence relation and the lens
laws are taken modulo equivalence. This allows lenses to be simpler and more
flexible. For example, inessential details like whitespace and orderings of
fields can be ignored when writing the transformations between the formats.

>>>>

Oftentimes, the same data is stored in a variety different formats. For
example, \bibtex{} and EndNote both store bibliography data; spreadsheet data
can be represented as a CSV file or a TSV file; web APIs can return data as a
JSON or
XML object. When different applications manipulate data in different formats,
synchronizing the applications amounts to synchronizing their respective data
formats. One way to synchronize data formats is to use \emph{lenses}, \emph{i.e.,} programs that represent pairs of functions for translating data from
format A to B (the \emph{get} direction) and back again (the \emph{put}
direction).  These lenses, when defined using domain-specific
languages like Boomerang~\cite{boomerang} or
BiGul~\cite{DBLP:conf/pepm/KoZH16}, typically come with a variety of guarantees,
the \emph{lens laws}, that help ensure put and get functions are effective
inverses of one another.
%comprise such synchronization functions, and the ``lens
%laws'' provide restrictions on how these synchronization functions interoperate%.
%Bidirectional languages like Boomerang~\cite{boomerang} and
%BiGul~\cite{DBLP:conf/pepm/KoZH16} use a single expression to encode all of the
%lens's synchronization functions at the same time.

One way to classify lenses is by the laws they obey.  Amongst the most
restrictive lenses are the \emph{bijective lenses}.  These lenses demand
that get and put functions for a bijection, meaning no information
is lost when one translates from data format to another.  While this
strong guarantee can be useful in some contexts, it is more often the case
that some of the information in one or both of the data formats is superfluous.
For instance, one typically does not care to preserve the amount of
whitespace between characters.  In other circumstances, capitalization
or the order of fields in a record or items in a list may be unimportant.
%
%bijective lenses have
%strong correctness guarantees, they are often
%However, these lens laws are too restrictive for many important transformations.
In particular, these lens laws do not permit transformations that put the data
in canonical forms.
For instance, to return to our bibliography example, in \bibtex{},
an author can be specified as
``\cd{lastname, firstname}'' or as ``\cd{firstname lastname}''---it does not matter which of the
two formats is used.
%While it is important to allow authors
%to be input in both ways, we would like these authors to only be output as
%\cd{firstname lastname}, but the lens laws prevent such functions.
{\em Quotient lenses}~\cite{quotientlenses} loosen many of the restrictions
of bijective lenses by introducing an equivalence relation and considering
bijections modulo that relation.  Such quotient lenses make it possible
to align many more data sources than would otherwise be possible by
ignoring details that a programmer views as inessential.
%each of the data formats is quotiented by some equivalence relation and the lens
%laws are taken modulo equivalence.
%This allows lenses to be simpler and more
%flexible. For example, inessential details like whitespace and orderings of
%fields can be ignored when writing the transformations between the formats.
